Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-10-03 packages/spec-haskell/yellowpaper/Paper1.tex
Agreement Framework

\section{Hierarchy of Agreements}

>>>>

The main tasks of the implementer of \textit{denotational payment primitives} are:

\begin{itemize}
    \item preserve the program correctness (and if possible, with proof of equivalence),
    \item optimize for efficient computation,
    \item and provide a software interface for using it.
\end{itemize}

The protocol introduces \textit{agreement framework} to address the optimization needs and to
provide a consistent software interface called ``agreement'' with agreement laws for reasoning about
the correctness.

\paragraph{Monetary Unit Data}

Recall that all monetary units have a monetary value function: $$ \nu :: u \rightarrow \mathbb{N}$$.

Agreement framework defines a concept called \textit{monetary unit data}, and each monetary unit has
a set of them:

\begin{code}
class ( SuperfluidCoreTypes sft
      ) => MonetaryUnitDataClass mud sft | mud -> sft where
    -- | π function - balance provided (hear: π) by the monetary unit data.
    balanceProvided
        :: forall t rtb.
           -- indexed type aliases
           ( t ~ SFT_TS sft
           , rtb ~ SFT_RTB sft
           )
        => mud -> t -> rtb
\end{code}

The $\pi$ function produces \textit{real time balance} at a particular time for \textit{monetary
unit data}. How a monetary unit produces monetary value from its set of monetary unit data is
described in the ``hierarchy of agreements'' section.

\paragraph{Agreement Contract}

Recall how a \textit{payment primitive} generator should look like:

Shared context $ctx$ is clearly the only data can be used by optimization, since it is updated each
time a primitive is generated. In the agreement framework, a data type \textit{agreement contract}
in the shared context is to fulfill this role:

\begin{code}
-- | Agreement contract type class.
class ( SuperfluidCoreTypes sft
      , Default ac
      , MonetaryUnitDataClass ac sft
      , Default (AgreementOperationOutput ac)
      , Traversable (AgreementOperationOutputF ac) -- <= Foldable Functor
      ) => AgreementContract ac sft | ac -> sft where

    -- | ω function - apply agreement operation ~ao~ (hear: ω) to the agreement
    --                operation data ~ac~ to get a tuple of:
    --
    --   1. An updated ~ac'~.
    --
    --   2. A functorful delta of agreement monetary unit data ~mudsΔ~, which
    --      then can be monoid-appended to existing
    --      ~mudΔ~.  This is what can make an agreement scalable.
    applyAgreementOperation
        :: forall t ao aoo.
           -- indexed type aliases
           ( t ~ SFT_TS sft
           , ao ~ AgreementOperation ac
           , aoo ~ AgreementOperationOutput ac
           )
        => ac -> ao -> t -> (ac, aoo)
    data family AgreementOperation ac :: Type
    data family AgreementOperationOutputF ac :: Type -> Type
    type family AgreementOperationOutput ac = (mmuds :: Type) | mmuds -> ac
\end{code}

The $\omega$ function ($applyAgreementOperation$) is the $genPrim$ in the agreement framework. It
takes in a agreement contract, an operation onto it and the current time; then it spits out an
update of the agreement contract and a functorful of new monoidal monetary unit data.

Here is a visualization of ``machinery'' of the $\omega$ function:

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{../assets/agreement-contract-omega-function.png}
    \caption{Agreement Contract $\omega$ Function}
    \label{fig:ac-omega}
\end{figure}

It also illustrates that it is expected the process should respect the \textit{law of conservation
of value} mandated by the \textit{restricted money distribution} model.

It is also important to note that $AgreementContract$ itself is also a $MonetaryUnitDataClass$, but
it is not a monoid, hence it is used to replace the previous agreement contract data. This has
optimization implications, as in that if agreement contracts do not produce zero balance then they
must be included in the $\nu$ function, that can make the $\nu$ function $O(N)$ to the number of
agreement contracts a monetary unit is associated with.

It may also have been self evident that agreement contract is a analogy to the fact that it encodes
``ongoing relationships'' like legal contracts do between bearers.

\subsection{Hierarchy of Agreements}
